---
title: "Equifax - Event Study"
author: "Fredrik Larsson"
date: "June 17, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Abstract
Conduct an event study to study whether the data breach of Equifax 2017 involved insider trading, abnormal market behavior (deviating from Efficient Market Hypothesis), and structural changes in the stock price.

### Packages
1. quantmod
2. dygraphs
3. forecast
4. vars
5. tseries

```{r, include=FALSE}
library(quantmod)
library(dygraphs)
library(forecast)
library(vars)
library(tseries)
```

### Import and Fix Data
After the announcement of the data breach, multiple noticeable events occurred as a direct result (arguably both positive and negative in respect to stock return). For example, the retirement of the CEO could be viewed as positive event, as it showcased that the board of Equifax had zero tolereance toward insider trading & corporate negligence.

Therefore, I will seperate events into two groups: positive news and negative news. 

$$Formula$$
$$R_t = \ln \frac {P_t}{P_{t-1}} = \ln P_t - \ln P_{t-1}$$

```{r}
#Stock Data
start.date <- "2016-12-01"
end.date <- "2018-06-01"
periodicity <- "daily"

stock.list <- c("EFX")

firm_stocks <- getSymbols.yahoo(Symbols = stock.list,
                                env = .GlobalEnv,
                                return.class = "xts",
                                index.class = "Date",
                                from = start.date,
                                to = end.date,
                                periodicity = periodicity)


#News variables - dates & explanation
announcement.dates <- c("2017-09-07",
                        "2017-09-11",#Announcement made a non-trading day. 
                        "2017-09-15",
                        "2017-09-26",
                        "2017-10-02")

announcement.explain <- c("Announces breach",
                          "Tweets phishing site",
                          "CSO & CIO retire",
                          "CEO retires",
                          "Announces more affected")

#Dummy variables - if news occured after market closure, I moved it forward to the next trading day.
negative.D <- c("2017-09-08",#Happened 2017-09-07
                "2017-09-11",#Happened 2017-09-10
                "2017-10-02")

positive.D <- c("2017-09-18",#Announcement made after trading hours on 2017-09-15
                "2017-09-26")

#Create consolidated data frame with stock returns (Stock price being non-stationry (mean and variance))
df <- cbind.xts(diff(log(EFX$EFX.Adjusted)),
                as.integer(as.character(index(EFX)) %in% as.character(negative.D)),
                as.integer(as.character(index(EFX)) %in% as.character(positive.D)))

names(df) <- c("Return", "Negative.D", "Positive.D")
head(df); subset(df, df$Negative.D == 1);subset(df, df$Positive.D == 1)#Print sample rows. 
```

#### Graph Stock Price, Stock Volume, and Event Dates
To get a better understanding of stock behavor during this series of events.
```{r}
dygraph(EFX[,5:6], xlab = "Date", ylab = "Stock Price", main = "Equifax Stock Price") %>%
  dyRangeSelector(dateWindow = c("2017-08-15", "2017-10-31")) %>%
  dyAxis("y", label = "Price", independentTicks = TRUE) %>%
  dyAxis("y2", label = "Volume", independentTicks = TRUE, valueRange = c(0, (10**8))) %>%
  dySeries("EFX.Volume", axis = ("y2")) %>%
  dyRoller(rollPeriod = 1) %>%
  dyEvent(announcement.dates, label = announcement.explain, col = "orange", labelLoc = "top") %>%
  dyOptions(drawGrid = FALSE)%>%
  dyLegend(width = 400)
```

### Create Benchmark Model: Fama-French 5 Factor Model
With help of a benchmark model (which will be based on data from a period prior to the first even), we can calculate abnormal returns and their respective t-values to detect any potential insider trading or deviation from EMH.

$$Formulas$$
$$R_{EFX} - R_{rf} = \alpha + \beta_1 MRP + \beta_2 SMB + \beta_3 HML + \beta_4 RMW + \beta_5 CMA + \epsilon$$
To estimate the coefficients of this model, I will be using an OLS regression (I assume that all 5/6 assumptions are met). 

$$R_{Abnormal} = R_{Actual} - R_{Expected}$$
$$t.value = \frac{R_{Abnormal}}{RSE}$$

```{r}
#Import fama-french variables, downloaded from Fama's website. *Excel sheet is adjusted. 
FF.factors <- read.csv("C:/Users/fredr/Desktop/webpage final/larssonft.github.io-master/F-F_Research_Data_5_Factors_2x3_daily.CSV")
ff.df <- data.frame(df$Return, FF.factors[,2:7]/100); ff.df <- as.xts(ff.df)#/100 to convert to decimal form.

#Seperate data into training and testing sets so that the model does not account for large unexpected returns. 
n.days <- 15#How many days before the first announcement should we study (important for insider trading)
train.df <- subset(ff.df, as.character(index(ff.df)) < as.Date(announcement.dates[1]) - n.days)
test.df <- subset(ff.df, as.character(index(ff.df)) > as.Date(announcement.dates[1]) - (n.days+1) &
                    as.character(index(ff.df)) < as.Date(announcement.dates[5]) + n.days)

#Create OLS model. To meet the 5/6 underlying assumptions for OLS regression, I will use returns (Stationarity).
ff <- lm((Return - RF) ~ Mkt.RF + SMB + HML + RMW + CMA, data = train.df)
summary(ff)#Only MRP is significantly different from 0; this model can only explain ~26% of variance in EFX Return.

#Apply the estimated coefficients/model on the testing data to detect abnormal returns.
comp.df <- cbind.xts(test.df$Return, #Actual Returns
                     predict(ff, test.df) + test.df$RF, #Predicted Return (FF-5 Factor)
                     test.df$Return -  predict(ff, test.df) + test.df$RF, #Abnormal Returns
                     (test.df$Return -  predict(ff, test.df) + test.df$RF)/summary(ff)$sigma)#t-value Abnormal Return

names(comp.df) <- c("Actual.R", "Expected.R", "Abnormal.R", "t.value")
t.value <- 1.96#Alpha = .95

#Extract dates with abnormal returns that are significantly different from 0 (i.e. absoulte t-value > 1.96).
abnormal.dates <- subset(index(comp.df), abs(comp.df$t.value) > t.value) 

#Graph stock price, event dates, and dates with abnormal returns. 
dygraph(comp.df[,1:2], xlab = "Date", ylab = "Return", main = "Actual vs. Expected Returns") %>%
  dyRangeSelector(dateWindow = c("2017-09-05", "2017-10-03")) %>%
  dyRoller(rollPeriod = 1) %>%
  dyEvent(announcement.dates, label = announcement.explain, col = "orange", labelLoc = "bottom") %>%
  dyEvent(abnormal.dates, label = c("Abnormal", 2:length(abnormal.dates)), col = "purple", labelLoc = "bottom") %>%
  dyOptions(drawGrid = FALSE)%>%
  dyLegend(width = 400)
```

Looking at the dates with abnormal returns, we failed to detect any potential insider trading. The CEO was, however, charged with insider trading as he unloaded his position in Equifax (~5M) prior to announcing the breach. I suspect that ~5M is insufficiently large to create any distortion in the stock price.

On the other hand, it appears that EFX deviated from Efficient Market Hypothesis, as the market failed to react instantly to new information.

### Fama-French 5 Factor Model + Dummy Variables
Evidentially, the data breach had a significantly negative impact on the stock price. By using dummy variables, we can try to quantify the actual impact on the days when new information arises. In this scenario, we don't need to seperate the data into training and testing sets, as we only want to measure the impact of the dummy variable. 

$$Formula$$
$$R_{EFX} - R_{rf} = \alpha + \beta_1 MRP + \beta_2 SMB + \beta_3 HML + \beta_4 RMW + \beta_5 CMA + \phi_1 D_N + \phi_2 D_P + \epsilon$$
```{r}
ff.df <- data.frame(ff.df, df$Negative.D, df$Positive.D); ff.df <- as.xts(ff.df)#Add dummy variables

ff.d <- lm(Return ~ Mkt.RF + SMB + HML + RMW + CMA + Negative.D + Positive.D, data = ff.df)
summary(ff.d)
```

Based on the estimated coefficients and their t-values, we can exepect that a similar negative event (and subsequent negative events) will generate a -7.5% return the same day. Subsequent positive events, however, are not significantly different from 0%, and thus cannot be estimated. 

### Impulse Response Function + Vector Autoregression
Until this point, we have a rough estimate what will happen to the stock price on the dates of announcement. Nevertheless, our analysis fails to consider the long-run outcome. With help of IRF and VAR, we will try to estimate any structural changes to the stock price.  

Firstly, I will analyze the stock price, and secondly, I will look at stock returns.

In reality, one should be careful with non-stationary variables (in this case, only return is a stationary variable).

<br/><br/>

#### Stock Price
Looking at the IRF graph, it appears to be a structural change to the stock price of about -$38 in the long-run.  
```{r}
df.price <- cbind.xts(EFX$EFX.Adjusted, df$Negative.D, df$Positive.D); names(df.price) <- c("Price", "N", "P")

#VARselect(df.price, lag.max = 25)#16 Lags is the optimal model (Lowest AIC)
VAR.price <- VAR(df.price, type = "const", p = 16)#p = # of lags

IRF.price <- irf(VAR.price, impulse = c("N", "P"), cumulative = TRUE, n.ahead = 100)

IRF.price.change <- data.frame(IRF.price$irf$N[,1], IRF.price$Lower$N[,1], IRF.price$Upper$N[,1],
                               IRF.price$irf$P[,1], IRF.price$Lower$P[,1], IRF.price$Upper$P[,1])
names(IRF.price.change) <- c("N.Price", "N.L", "N.H", "P.Price", "P.L", "P.H")

IRF.price.change <- ts(IRF.price.change, start = 0, end = (nrow(IRF.price.change)-1))

#For dygraph, I used a time-series (ts) object as xts requires specific dates. 
dygraph(IRF.price.change, main = "Cumulative Change in Price", xlab = "Days after Event", ylab = "Price") %>%
  dySeries(c("N.L", "N.Price", "N.H"), label = "Negative News", col = "red") %>%
  dySeries(c("P.L", "P.Price", "P.H"), label = "Positive News", col = "green") %>%
  dyRangeSelector() %>%
  dyLegend(width = 400)
```

#### Stock Return
The negative news appear to have an effect on the stock price up to 40 days after the announcement. Of course, this study did not cover all new events surrounding Equifax during this time.  
```{r}
df.return <- cbind.xts(df$Return, df$Negative.D, df$Positive.D); names(df.return) <- c("Return", "N", "P")
df.return <- na.omit(df.return)

#VARselect(df.return, lag.max = 25)#16 Lags is the optimal model (Lowest AIC)
VAR.return <- VAR(df.return, type = "const", p = 16)#p = # of lags

IRF.return <- irf(VAR.return, impulse = c("N", "P"), cumulative = FALSE, n.ahead = 50)

IRF.return.change <- data.frame(IRF.return$irf$N[,1], IRF.return$Lower$N[,1], IRF.return$Upper$N[,1],
                               IRF.return$irf$P[,1], IRF.return$Lower$P[,1], IRF.return$Upper$P[,1])

names(IRF.return.change) <- c("N.Return", "N.L", "N.H", "P.Return", "P.L", "P.H")

IRF.return.change <- ts(IRF.return.change, start = 0, end = (nrow(IRF.return.change)-1))

#For dygraph, I used a time-series (ts) object as xts requires specific dates. 
dygraph(IRF.return.change, main = "Return after Shock", xlab = "Days after Event", ylab = "Return") %>%
  dySeries(c("N.L", "N.Return", "N.H"), label = "Negative News", col = "red") %>%
  dySeries(c("P.L", "P.Return", "P.H"), label = "Positive News", col = "green") %>%
  dyRangeSelector() %>%
  dyLegend(width = 400)
```



