<br/><br/>
Hello! 

My name is Fredrik Larsson, and I am currently pursuing a MS in Finance at USC Marshall School of Business. Please see a list of available projects below or in the bar menu above. My full CV is available [here](CV/CV.pdf).

<br/><br/>

```{r, echo=FALSE}
library(htmltools)
thumbnail <- function(title, img, href, caption = TRUE) {#https://garrettgman.github.io/rmarkdown/rmarkdown_websites.html
  div(class = "col-sm-4",
      a(class = "thumbnail", title = title, href = href,
        img(src = img),
        div(class = ifelse(caption, "caption", ""),
          ifelse(caption, title, "")
        )
      )
  )
}
```

```{r, echo=FALSE}
thumbnail("Portfolio Optimization", "portfolio_optimization_files/figure-html/unnamed-chunk-4-1.png", "https://larssonft.github.io/portfolio_optimization.html")
```
Considering the allocation of capital in a portfolio of assets is highly important to maximize returns in respect to risk - this can be both systematic and idiosyncratic risk. In this project, I will utilize matrix algebra to find the tangency portfolio, using $\sum_{}^{-1}\times Risk Premium$. However, I assume that volatility, asset covariance, and returns are the same as in the past (I do not recommend this approach when investing real money). Then, I will compare this sophisticated portfolio statistically to an equal weight portfolio, SPY (S&P500 ETF), and VV (Vanguard Large Cap Portfolio) in 2019. It looks at the relative performance and how much the investor would have been compensated based risk levels. The conclusion is that a historical Sharpe Ratio is an insufficient measure to determine portfolio weights, although it is much more mathematical and computational intense.

<br/><br/>
<br/><br/>



```{r, echo = FALSE}
thumbnail("Event Study", "images/event_study.PNG", "https://larssonft.github.io/event_study.html")
```
Event Study looking into how the data breach of Equifax 2017 affect the stock price. Various methods have been use in order to determine direct and long-term consequences. 

<br/><br/>
<br/><br/>
<br/><br/>
<br/><br/>
<br/><br/>


```{r, echo=FALSE}
thumbnail("Fama French - 3 Factor Model", "fama_french_files/figure-html/unnamed-chunk-4-1.png", "https://larssonft.github.io/fama_french.html")
```
Since the uprise of CAPM, academia has tried to find new models to explain stocks and portfolio movements. Some of the best to date are the Fama French 3-Factor and 5-Factor Models. Many researchers, however, have tried to add more factors, but it has been unsuccessful as more independent variables tend to punish the model statistically: lower $adj. R^2$. Now, researchers and industry professionals are trying to find more meaningful variables. 

The 3 factors are: MRP (Market Risk Premium), HML (High minus Low), and SML (Small minus Big). I utilized ETFs to capture the different categories of stocks rather than downloading the whole market. At the end, I lag all independent variables with 1 day to see if we can use Fama French to predict the price of tomorrow.












