---
title: "Portfolio Optimization"
author: "Fredrik Larsson"
date: "January 11, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Packages:
Quantmod; PerformanceAnalytics; PortfolioAnalytics; forecast; timeSeries
```{r, include=FALSE}
library(quantmod)
library(PerformanceAnalytics)
library(PortfolioAnalytics)
library(forecast)
library(timeSeries)
```

### Download Stock and Market Data
```{r}
stocks <- c("AMZN", "JPM", "KSS", "CMG", "JNJ", "^TNX", "SPY")
start.date <- "2016-01-01"
end.date <- "2018-12-31"

stock.data <- getSymbols.yahoo(Symbols = stocks,
                               env = .GlobalEnv,
                               return.class = "xts",
                               index.class = "Date",
                               from = start.date,
                               to = end.date,
                               periodicity = "monthly")

monthly.returns <- cbind.xts(ROC(AMZN$AMZN.Adjusted),
                             ROC(JPM$JPM.Adjusted),
                             ROC(KSS$KSS.Adjusted),
                             ROC(CMG$CMG.Adjusted),
                             ROC(JNJ$JNJ.Adjusted),
                             TNX$TNX.Adjusted/1200)#Convert to montly decimal data.

names(monthly.returns) <- stocks[1:6]; monthly.returns <- na.omit(monthly.returns)
head(round(monthly.returns,3))#Example data.
chart.CumReturns(monthly.returns, wealth.index = TRUE, legend.loc = "topleft")#Total Returns of each asset.
```

### Construct a Portfolio
I will construct and compare 2 different portfolios: Equal Weight and Tangency Portfolio.

```{r}
#Equal Weight: 100% / 5 = 20%.
equal.weights <- c(0.2, 0.2, 0.2, 0.2, 0.2)

#Global Minimum Variance


#Tangency Portfolio
risk.free.rate <- mean(monthly.returns$`^TNX`)#Risk Free Rate to be used.
cov.mat <- cov(monthly.returns[,1:5]); round(cov.mat,3)# Variance Covariance Matrix.

risk.premium <- matrix(c(mean(monthly.returns$AMZN), #RP = ri - rf.
                         mean(monthly.returns$JPM),
                         mean(monthly.returns$KSS),
                         mean(monthly.returns$CMG),
                         mean(monthly.returns$JNJ)), ncol = 1) - risk.free.rate#rf
risk.premium# Risk Premium Matrix

tangency.weights <- solve(cov.mat, risk.premium)/sum(solve(cov.mat, risk.premium))#Solve for tangency point.
colnames(tangency.weights) <- "Weights"; round(tangency.weights, 3)
```

### Graph Efficient Frontier
Graph Efficient frontier to see how the different portfolios lie relatively to each other. 
```{r}
#Generate Random Portfolio using PortfolioAnalytics package.
portfolio.generation <- portfolio.spec(stocks[1:5])#Add assets' names.
portfolio.generation <- add.constraint(portfolio.generation, type = "box", min = -1, max = 1)#Add constraints.
portfolio.generation <- add.constraint(portfolio = portfolio.generation, type = "full_investment")
random.portfolios <- random_portfolios(portfolio.generation, permutations = 2000, rp_method = "grid")

portfolio.returns <- random.portfolios%*%(risk.premium+risk.free.rate)
portfolio.std <- c()#Create empty list to fill with portfolio standard deviations.
for(i in (1:nrow(random.portfolios))){
  portfolio.std <- c(portfolio.std,(random.portfolios[i,]%*%cov.mat%*%matrix(c(random.portfolios[i,])))**(1/2))
}

equal.weights.std <- (equal.weights%*%cov.mat%*%matrix(c(equal.weights)))**(1/2)
equal.weights.return <- equal.weights%*%(risk.premium+risk.free.rate)

tangency.std <- (t(tangency.weights)%*%cov.mat%*%tangency.weights)**(1/2)
tangency.return <- t(tangency.weights)%*%(risk.premium+risk.free.rate)

plot(portfolio.std, portfolio.returns, col = "grey", xlab = "Std. Dev.", ylab = "Return", main = "Efficient Frontier", xlim = c(0,0.09), ylim = c(-.015, 0.03));grid(); points(c(0, equal.weights.std, tangency.std), c(risk.free.rate, equal.weights.return, tangency.return), col = c("red", "green", "purple")); points(diag(cov.mat)**(1/2), risk.premium+risk.free.rate, col = "blue"); text(c(0.01,0.02,0.05,0.07), c(0, 0.015, 0.027, 0.017), c("Risk Free Rate", "Equal Weights Portfolio", "Tangency Portfolio", "Individual Assets"), col = c("red", "green", "purple", "blue")); abline(risk.free.rate, ((tangency.return - risk.free.rate)/ tangency.std), col = "purple")# Graph function looks complicated as I added many features.
```

### Backtest how the portfolio performed relatively to SPY 2019M1 - 2019M11.
Need to download daily data - which will be data used from hereon. 
```{r}
start.date <- "2019-01-01"; end.date <- "2019-11-30"
stock.data <- getSymbols.yahoo(Symbols = stocks,
                           env = .GlobalEnv,
                           return.class = "xts",
                           index.class = "Date",
                           from = start.date,
                           to = end.date,
                           periodicity = "daily")

daily.returns <- cbind.xts(ROC(AMZN$AMZN.Adjusted),
                             ROC(JPM$JPM.Adjusted),
                             ROC(KSS$KSS.Adjusted),
                             ROC(CMG$CMG.Adjusted),
                             ROC(JNJ$JNJ.Adjusted),
                             ROC(SPY$SPY.Adjusted))   
daily.returns <- na.omit(daily.returns)#Remove any potential NAs.
names(daily.returns) <- c(stocks[1:5], stocks[7])

equal.weights.daily.return <- rowSums(daily.returns[,1:5]*0.2)
tangency.daily.return <- (daily.returns$AMZN*tangency.weights[1]
                          + daily.returns$JPM*tangency.weights[2]
                          + daily.returns$KSS*tangency.weights[3]
                          + daily.returns$CMG*tangency.weights[4]
                          + daily.returns$JNJ*tangency.weights[5])

portfolios.daily.return <- cbind.xts(equal.weights.daily.return, tangency.daily.return, daily.returns$SPY)
names(portfolios.daily.return) <- c("Equal Weight", "Tangency", "SPY")

head(round(portfolios.daily.return,3))#Example data.
chart.CumReturns(portfolios.daily.return, legend.loc = "topleft")#Graph cumulative return of each portfolio
#We can see that the tangency portfolio seriously underperform the market and only the euqal weight portfolio exceeded the market once in respect to returns. 
```

### Estimate CAPM model for both portfolios.
CAPM to be based on daily returns in 2019. 
```{r}
risk.free.daily.rate <- mean(na.omit(TNX$TNX.Adjusted)/(100*250))#250 trading days in 1 year

regression.equal.weight <- lm(portfolios.daily.return$`Equal Weight` ~ portfolios.daily.return$SPY,
                              data = portfolios.daily.return[, c("Equal Weight", "SPY")] -
                              risk.free.daily.rate)#Remove risk free rate from Port. and SPY returns.
summary(regression.equal.weight)#t-value is signifcant enough to reject the H0: beta = 0. However, the t-value for alpha is too small to conclude that Ha: a =/= 0.
confint(regression.equal.weight)#Confidence Interval.

regression.tangency <- lm(portfolios.daily.return$Tangency ~ portfolios.daily.return$SPY,
                              data = portfolios.daily.return[, c("Tangency", "SPY")] -
                              risk.free.daily.rate)
summary(regression.tangency)#Both t-values are signifcant enough to reject the H0: beta = 0 and H0: a = 0.
confint(regression.tangency)#Confidence Interval.
```

### Graph Security Market Line (SML).
Visually we can see that both portfolios' returns do not provide enough returns for the amount of systematic risk taken on: therefore, following a market portfolio would be better in most instances (unless you are trying to hedge.)
```{r}
plot(c(0,3), c(risk.free.daily.rate, mean(portfolios.daily.return$SPY)*3), type = "l", xlab = "Beta", ylab = "Return"); grid(); points(c(regression.equal.weight$coefficients[2], regression.tangency$coefficients[2]), c(mean(portfolios.daily.return$`Equal Weight`), mean(portfolios.daily.return$Tangency)), col = c("green", "purple"));text(c(1.1, 1.4), c(0.0006, 0.00015), c("Equal Weight", "Tangency"), col = c("green", "purple"))
```

### Calculate the CV-, Sharpe-, Treynor-, and Sortino Ratio.
Compare to a similarly diverse portfolio. For this comparison, I will use a large cap ETF from Vanguard: VV. 
```{r}
Vanguard <- getSymbols.yahoo(Symbols = c("VV"),
                           env = .GlobalEnv,
                           return.class = "xts",
                           index.class = "Date",
                           from = "2019-01-01",
                           to = "2019-11-30",
                           periodicity = "daily")

ratio.comparison.returns <- cbind.xts(portfolios.daily.return, na.omit(ROC(VV$VV.Adjusted)))#Combine data.

#USe functions from PerformanceAnalytics, except for CV Ratio.
#Create a matrix with all the ratios. Of course, one could have created a for loop or print each function individually. 
comparison <- matrix(c(SharpeRatio(ratio.comparison.returns$Equal.Weight, Rf = risk.free.daily.rate)[1],#Create a matrix with all results.
                       SharpeRatio(ratio.comparison.returns$Tangency, Rf = risk.free.daily.rate)[1],
                       SharpeRatio(ratio.comparison.returns$SPY, Rf = risk.free.daily.rate)[1],
                       SharpeRatio(ratio.comparison.returns$VV.Adjusted, Rf = risk.free.daily.rate)[1],
                       TreynorRatio(ratio.comparison.returns$Equal.Weight, ratio.comparison.returns$SPY, Rf = risk.free.daily.rate),
                       TreynorRatio(ratio.comparison.returns$Tangency, ratio.comparison.returns$SPY, Rf = risk.free.daily.rate),
                       TreynorRatio(ratio.comparison.returns$SPY, ratio.comparison.returns$SPY, Rf = risk.free.daily.rate),
                       TreynorRatio(ratio.comparison.returns$VV.Adjusted, ratio.comparison.returns$SPY, Rf = risk.free.daily.rate),
                       SortinoRatio(ratio.comparison.returns$Equal.Weight, MAR = 0),
                       SortinoRatio(ratio.comparison.returns$Tangency, MAR = 0),
                       SortinoRatio(ratio.comparison.returns$SPY, MAR = 0),
                       SortinoRatio(ratio.comparison.returns$VV.Adjusted, MAR = 0),
                       StdDev(ratio.comparison.returns$Equal.Weight)/mean(ratio.comparison.returns$Equal.Weight),
                       StdDev(ratio.comparison.returns$Tangency)/mean(ratio.comparison.returns$Tangency),
                       StdDev(ratio.comparison.returns$SPY)/mean(ratio.comparison.returns$SPY),
                       StdDev(ratio.comparison.returns$VV.Adjusted)/mean(ratio.comparison.returns$VV.Adjusted)), ncol = 4)

rownames(comparison) <- c("Sharpe Ratio", "Treynor Ratio", "Sortino Ratio", "CV Ratio")
colnames(comparison) <- c("Equal Weight", "Tangency", "SPY", "VV")
comparison
```

### Calculate Daily 2% VaR (Value at Risk) as percentage of mean return.
This calculation can easily be done by using the VaR() function supplied by the PerformanceAnalytics Package.
```{r}
VaR.comparison <- c()
VaR(ratio.comparison.returns, p = 0.02)

chart.Histogram(ratio.comparison.returns$Equal.Weight,#Plot Density of Returns.
                breaks = 50, methods = c("add.density", "add.normal", "add.rug", "add.risk"), p =0.98, main = "Daily VaR Equal Weight")

chart.Histogram(ratio.comparison.returns$Tangency,#Plot Density of Returns. One with more data points may want to use more breaks.
                breaks = 50, methods = c("add.density", "add.normal", "add.rug", "add.risk"), p =0.98, main = "Daily VaR Tangency")
```

### Graph Scatter Plot to visually find any significant outliers and trends.
Scatter Plot will have SPY on X-axis and the other portfolios on the Y-axis. Equal Weight Portfolio conforms quite nicely to the performance of SPY (this was expected due to a beta almost = 1).
```{r}
#Use timeSeries() function to move away from XTS object that has predetermined properties in plot().
plot(timeSeries(ratio.comparison.returns$SPY), timeSeries(ratio.comparison.returns$Equal.Weight), ylab = "Equal Weight Return", xlab = "S&P 500 Return", main = "Equal Weight vs. SP500"); grid(); abline(0,1, col = "red"); abline(0, regression.equal.weight$coefficients[2], col = "green"); text(c(-.022,0.027),c(-.01,.015), c("Equal Weight Beta Slope", "Market Beta Slope"), col = c("green", "red"))

plot(timeSeries(ratio.comparison.returns$SPY), timeSeries(ratio.comparison.returns$Tangency), ylab = "Tangency Return", xlab = "S&P 500 Return", main = "Tangency vs. SP500"); grid(); abline(0,1, col = "red"); abline(0, regression.tangency$coefficients[2], col = "green"); text(c(-.02,0.027),c(-.04,.015), c("Tangency Beta Slope", "Market Beta Slope"), col = c("green", "red"))
```

### For December, I want to hold a Portfolio with 0 Systematic Risk. 
Create Portfolio with 0 Systematic Risk using data of Portfolios and SPY 2019M1 - 2019M11. Easiest way to get a portfolio with 0 systematic risk is to take a short position in SPY - contrary, you may take a long position in SPY and then short the portfolio (+ = - and - = +). However, someone who wants to devote their time to find a portfolio with 0 systematic risk should search for quant-strategies, as you want to add value - otherwise, just invest in risk free assets.  

We can see that we somewhat succeeded in reducing the systematic risk, given that our beta for each portfolio is close to 0 with a low t-value, suggesting that we fail to reject the H0: beta = 0. 
```{r}
start.date <- "2019-12-01"; end.date <- "2019-12-31"
stock.data <- getSymbols.yahoo(Symbols = stocks,
                           env = .GlobalEnv,
                           return.class = "xts",
                           index.class = "Date",
                           from = start.date,
                           to = end.date,
                           periodicity = "daily")

daily.returns <- cbind.xts(ROC(AMZN$AMZN.Adjusted),
                             ROC(JPM$JPM.Adjusted),
                             ROC(KSS$KSS.Adjusted),
                             ROC(CMG$CMG.Adjusted),
                             ROC(JNJ$JNJ.Adjusted),
                             ROC(SPY$SPY.Adjusted))

regression.equal.weight$coefficients[2]/1

equal.weights.daily.return <- rowSums(daily.returns[,1:5]*0.2)-daily.returns$SPY.Adjusted*(regression.equal.weight$coefficients[2])
# - = Short Position.
tangency.daily.return <- (daily.returns$AMZN*tangency.weights[1]
                          + daily.returns$JPM*tangency.weights[2]
                          + daily.returns$KSS*tangency.weights[3]
                          + daily.returns$CMG*tangency.weights[4]
                          + daily.returns$JNJ*tangency.weights[5]) - daily.returns$SPY.Adjusted*(regression.tangency$coefficients[2])

portfolios.daily.return <- cbind.xts(equal.weights.daily.return, tangency.daily.return, daily.returns$SPY)
names(portfolios.daily.return) <- c("Equal Weight", "Tangency", "SPY"); portfolios.daily.return <- na.omit(portfolios.daily.return)
head(portfolios.daily.return)

chart.CumReturns(portfolios.daily.return, legend.loc = "topleft", main = "Cumulative Return")
#Still underperformed the S&P500, but that was expected if S&P500 gained value during 2019M12. 

plot(portfolios.daily.return, legend.loc = "topleft", main = "Daily Returns")

regression.equal.weight <- lm(portfolios.daily.return$`Equal Weight` ~ portfolios.daily.return$SPY,
                              data = portfolios.daily.return[, c("Equal Weight", "SPY")] -
                              risk.free.daily.rate)#Remove risk free rate from Port. and SPY returns.

regression.tangency <- lm(portfolios.daily.return$Tangency ~ portfolios.daily.return$SPY,
                              data = portfolios.daily.return[, c("Tangency", "SPY")] -
                              risk.free.daily.rate)#Remove risk free rate from Port. and SPY returns.

summary(regression.equal.weight); summary(regression.tangency)
```







